\chapter*{Acknowledgements}Foremost, I would like to express my sincere gratitude to my advisor Prof.Sharat Chandran (CSE, IITB) and Prof. Michael Brown (School of Computing, NUSSingapore) for the continuous support in my Ph.D study and related research,for their patience, motivation, and immense knowledge.Besides my advisor, I would like to thank RPC committee members  Prof. ParagChaudhuri, and Prof. Ajit Rajwade for their insightful comments andencouragement, and also for the hard questions during APS which incented me towiden my research from various perspectives. I am also grateful to Prof. SuyashAwate (CSE, IIT Bombay) for his valuable guidance and support. I am very muchthankful to Prof. B. K. Chakravarthy (IDC, IIT Bombay), for giving permissionto use the infrastructure required for my experiments. I would also like tothank Prof. David Hsu (School of Computing, NUS Singapore) for allowing me touse the robotics lab during my research stay in Singapore.I take this opportunity to thank Mr. Anirudh Vemula for his help in pathplanning work. I am also thankful to Mr.Sona Praneeth Akula for his help in multiple planesdetection work. I am indebted to both of them for their help in data collectionfor the project. I thank my fellow labmates Dilip Prasad (Singapore), Lan Ziquan (Singapore),and ViGIL labmates: Ravi Chalasani, Abhishek Chakraborty,  Saurabh Shigwan,Preeti Gopal, and R. Thyagarajan, for the stimulating discussions, for thesleepless nights we were working together before deadlines, and for all the funwe had in the last four years. I would also like to thank office staff from CSEdepartment for their support. Last but not the least, I would like to thank my parents, my in-laws, my son andmy sisters for supporting me spiritually throughout writing this thesis and my life in general. My personal thanks to Mr. Shirish Joshi and Mrs. Puja Joshifor their valuable help during my stay in Singapore. Finally, I would like tomention my wife, Gauri, whose patience with me never wore out. It would not bean exaggeration if I say that I will owe my degree to her.